---
title: "Power and such"
author: "Pat Byrne"
date: "5/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Clear global environment
rm(list = ls()) 
# Load packages
library(tidyverse)
library(DeclareDesign)
library(ggthemes)
library(truncnorm)
library(kableExtra)
set.seed(666)
```

```{r}
# Reading in data table from https://www.citypopulation.de/en/peru/puno/
puno.data <- read_csv('Puno_localities.csv')
```

```{r}
total.pop <- sum(puno.data$Population)

df.sum <- puno.data %>% 
  filter(
    Status %in% c('Locality')
  ) %>% 
  group_by(Province,Status) %>% 
  summarize(
    pop_pct = sum(Population)/total.pop,
    num = n()
  )

df.province <- puno.data %>% 
  group_by(Province) %>% 
  summarize(
    pop = sum(Population),
    num = n()
  )


df.affected <- puno.data %>%
  filter(
    Province %in% c('San Antonio de Putina', 'Melgar', 'Lampa', 'El Collao', 'Azangaro')
  ) %>%
  group_by(Province) %>%
  summarize(
    num = n()
  )

# Total number of localities n the 5 provinces most affected 
n.affected.localities <- as.numeric(sum(df.affected$num))

# Total livestock deaths in the 5 provinces most affected vs total livestock deaths across Puno
affected.pct.livestock = (233482 + 176241 + 132860 + 68291 + 54825)/694953
```

```{r}
# Reading in a list of districts in each of the 5 most affected provinces obtained from Wikipedia
sample.districts <- read_csv('Puno_affected_districts.csv')
districts <- sample.districts$District
```

```{r}
# Calculating typical livestock mortality percentage (numbers from alpaca_death.xlsx)
mean.mortality <- mean(c(1.58, 5.94, 1.12, 1.48, 4.47, 13.78, 0.84, 8.20, 2.71, 0.59))
sd.mortality <- sd(c(1.58, 5.94, 1.12, 1.48, 4.47, 13.78, 0.84, 8.20, 2.71, 0.59))

```


# Power Analysis by Locality

##declare_population()
```{r population, echo=TRUE}
population <- declare_population(
  districts = add_level(N=589, 
    livestock_mortality_pct = rgamma(n=N,  shape = 2, rate = (2/mean.mortality)),
    u = rnorm(n=N, mean = 0, sd = sd.mortality)
    )
  )

pop <- population()
```

##declare_potential_outcomes()
```{r po, echo=TRUE}
mortality_reduction = -1

potential_outcomes <- 
  declare_potential_outcomes(
    Y_D_0 = livestock_mortality_pct + u,
    Y_D_1 = livestock_mortality_pct + u + mortality_reduction)

po <- potential_outcomes(pop)

# We are proposing that the unit-level treatment effect is mortality_reduction in % reduction of livestock that die from extreme frost.
```

##declare_sampling()
```{r sample, echo=TRUE}
sampling <- declare_sampling(n = 197)
sam <- sampling(po)
```

##declare_assignment()
```{r assign, echo=TRUE}
assigning <- declare_assignment(m = round(nrow(sam)/2),
                  assignment_variable="D")
assigned <- assigning(sam)

# We're using DeclareDesign to randomly assign districts to treatment and control with a probability of 0.5. The variable 'm' denotes the number of units that declare_assignment() should place in each treatment condition, which we are asking to be equal to half of the sample per treatment condition (nrow(sam)/2).

```

```{r}
ggplot(data=assigned, aes(x=as.factor(D), y=livestock_mortality_pct)) +
geom_violin(aes(fill=as.factor(D), color=as.factor(D))) +
theme_minimal(base_size = 24) + xlab("Assignment")

# Random assignment should, in expectation, produce treatment and control groups that are statistically-identical on all observed and unobserved features.
```


##declare_reveal()
```{r reveal, echo=TRUE}
revealing <- declare_reveal(assignment_variables=D)
```

##declare_estimand()
```{r estimand, echo=TRUE}
estimand <- declare_estimand(ATE = mortality_reduction)
estimand(po)
```

##declare_estimator()
```{r estimator, echo=TRUE}
dim <- declare_estimator(Y ~ D, estimand = estimand,  
          model =  difference_in_means, label = "DIM") 
# Difference-in-means (Mean of treatment group - mean of control group)

did <- declare_estimator(Y - livestock_mortality_pct ~ D, 
                         estimand = estimand,  
          model =  difference_in_means, label = "DID") 
# Difference-in-differences ([Mean of treatment group @ endline - Mean of treatment group @ baseline] - [Mean of control group @ endline - mean of control group @ baseline])
```


##declare_design()
```{r design, echo=TRUE}
design <- population + potential_outcomes + sampling +
          assigning + revealing + estimand + dim + did
```


##diagnose_design()
```{r diagnosis, cache=TRUE}
# diagnosis <- diagnose_design(design, sims=1000)
# diagnosis$diagnosands_df[,c(1,3,5,9,11)] %>%
#   kable()
```


```{r}
designs <- redesign(design, mortality_reduction=seq(from=-3, to=0, by=0.05))

diagnoses <- diagnose_design(designs, sims=300)

diagnoses$diagnosands_df %>%
  filter(estimator_label=="DID") %>%
  ggplot(aes(x=mortality_reduction,y=power)) +
  geom_line() +
  geom_hline(yintercept=0.8,
             linetype="dashed", 
             color = "red") +
  labs(x='Estimated Percentage Point Reduction in Livestock Mortality (%)',
       y = 'Power',
       title = 'Power vs Potential Average Treatment Effect, Locality-Level Treatment',
       subtitle = 'Estimator = Difference in Differences') + 
  theme_tufte()
```







## Looking under the hood, DIM

```{r underhood-dim, height=6}
# sim.out <- diagnosis$simulations
# hist(sim.out$estimate[sim.out$estimator_label == "DIM"],
#      main="Randomization Distribution",
#      xlab="Estimates in Realized Experiments",
#      xlim=c(-10,10), cex=24)
# abline(v=-4, lwd=3, col="red")
```

## Looking under the hood, DID

```{r underhood-did, height=6}
# sim.out <- diagnosis$simulations
# hist(sim.out$estimate[sim.out$estimator_label == "DID"],
#      main="Randomization Distribution",
#      xlab="Estimates in Realized Experiments",
#      xlim=c(-10,10), cex=24)
# abline(v=-4, lwd=3, col="red")
```



# Power Analysis by District

##declare_population()
```{r population, echo=TRUE}
population <- declare_population(
  dstrcts = add_level(N=107, 
    livestock_mortality_pct = rgamma(n=N, shape = 2, rate = (2/mean.mortality)),
    u = rnorm(n=N, mean = 0, sd = sd.mortality)
    )
  )

pop <- population()
```


##declare_potential_outcomes()
```{r po, echo=TRUE}
mortality_reduction = -1

potential_outcomes <- 
  declare_potential_outcomes(
    Y_D_0 = livestock_mortality_pct + u,
    Y_D_1 = livestock_mortality_pct + u + mortality_reduction)

po <- potential_outcomes(pop)

# We are proposing that the unit-level treatment effect is mortality_reduction in % reduction of livestock that die from extreme frost.
```

##declare_sampling()
```{r sample, echo=TRUE}
sampling <- declare_sampling(n = 54)
sam <- sampling(po)
```

##declare_assignment()
```{r assign, echo=TRUE}
assigning <- declare_assignment(m = (nrow(sam)/2),
                  assignment_variable="D")
assigned <- assigning(sam)

# We're using DeclareDesign to randomly assign districts to treatment and control with a probability of 0.5. The variable 'm' denotes the number of units that declare_assignment() should place in each treatment condition, which we are asking to be equal to half of the sample per treatment condition (nrow(sam)/2).

```

```{r}
ggplot(data=assigned, aes(x=as.factor(D), y=livestock_mortality_pct)) +
geom_violin(aes(fill=as.factor(D), color=as.factor(D))) +
theme_minimal(base_size = 24) + xlab("Assignment")

# Random assignment should, in expectation, produce treatment and control groups that are statistically-identical on all observed and unobserved features.
```


##declare_reveal()
```{r reveal, echo=TRUE}
revealing <- declare_reveal(assignment_variables=D)
```

##declare_estimand()
```{r estimand, echo=TRUE}
estimand <- declare_estimand(ATE = mortality_reduction)
estimand(po)
```

##declare_estimator()
```{r estimator, echo=TRUE}
dim <- declare_estimator(Y ~ D, estimand = estimand,  
          model =  difference_in_means, label = "DIM") 
# Difference-in-means (Mean of treatment group - mean of control group)

did <- declare_estimator(Y - livestock_mortality_pct ~ D, 
                         estimand = estimand,  
          model =  difference_in_means, label = "DID") 
# Difference-in-differences ([Mean of treatment group @ endline - Mean of treatment group @ baseline] - [Mean of control group @ endline - mean of control group @ baseline])
```


##declare_design()
```{r design, echo=TRUE}
design <- population + potential_outcomes + sampling +
          assigning + revealing + estimand + dim + did
```


##diagnose_design()
```{r diagnosis, cache=TRUE}
# diagnosis <- diagnose_design(design, sims=500)
# diagnosis$diagnosands_df[,c(1,3,5,9,11)] %>%
#   kable()
```

```{r}
designs <- redesign(design, mortality_reduction=seq(from=-3, to=0, by=0.05))

diagnoses <- diagnose_design(designs, sims=300)

diagnoses$diagnosands_df %>%
  filter(estimator_label=="DID") %>%
  ggplot(aes(x=mortality_reduction,y=power)) +
  geom_line() +
  geom_hline(yintercept=0.8,
             linetype="dashed", 
             color = "red") +
  labs(x='Estimated Percentage Point Reduction in Livestock Mortality (%)',
       y = 'Power',
       title = 'Power vs Average Treatment Effect, District-Level Treatment in Top 6 Provinces',
       subtitle = 'Estimand = Difference in Differences, Sample Size = 54') +
  theme_tufte()
```


## Looking under the hood, DIM

```{r underhood-dim, height=6}
# sim.out <- diagnosis$simulations
# hist(sim.out$estimate[sim.out$estimator_label == "DIM"],
#      main="Randomization Distribution",
#      xlab="Estimates in Realized Experiments",
#      xlim=c(-10,10), cex=24)
# abline(v=mortality_reduction, lwd=3, col="red")
```

## Looking under the hood, DID

```{r underhood-did, height=6}
# sim.out <- diagnosis$simulations
# hist(sim.out$estimate[sim.out$estimator_label == "DID"],
#      main="Randomization Distribution",
#      xlab="Estimates in Realized Experiments",
#      xlim=c(-10,10), cex=24)
# abline(v=mortality_reduction, lwd=3, col="red")
```

